---
title: "Quasi Monte Carlo"
output: html_document
---

The first example of a QMC is the van Der corput sequence.

#Van der Corput
```{r, include=FALSE}
library(randtoolbox)
library(ggplot2)
library(GGally)
```
Let's generate a 10 point van der Corput sequence.
```{r}
npoints = 10
qmc <- data.frame(j = 1:npoints, x = halton(npoints))
ggplot(qmc,aes(x=x,y=j)) + geom_point(size=4) + scale_y_continuous('Sample Number')
ggplot(qmc,aes(x=j,y=x)) + geom_line() + geom_point(size=4) + scale_x_continuous('Sample Number')
```

#Halton Sequence
Now a Halton Sequence in 2-D
```{r}
npoints = 50
xy = halton(npoints,dim=2)
qmc <- data.frame(j = 1:npoints)
qmc[,c("x","y")] = xy
ggplot(qmc,aes(x=x,y=y)) + geom_point(size=4)
```

The warts begin to appear in a Halton Sequence in 6-D
```{r}
npoints = 50
xy = halton(npoints,dim=6)
qmc <- data.frame(j = 1:npoints)
qmc[,2:7] = xy
ggpairs(qmc,columns=2:7)
```

15-D is just plain bad.
```{r}
npoints = 50
xy = halton(npoints,dim=15)
qmc <- data.frame(j = 1:npoints)
qmc[,2:16] = xy
ggpairs(qmc,columns=10:15)
```

#Sobol Sequnce
In 2-D the Sobol sequence is similar to Halton.
```{r}
npoints = 50
xy = sobol(npoints,dim=2)
qmc <- data.frame(j = 1:npoints)
qmc[,c("x","y")] = xy
ggplot(qmc,aes(x=x,y=y)) + geom_point(size=4)
```

In 6-D the Sobol appears to be somewhat better than Halton, but still imperfect.
```{r}
npoints = 50
xy = sobol(npoints,dim=6)
qmc <- data.frame(j = 1:npoints)
qmc[,2:7] = xy
ggpairs(qmc,columns=2:7)
```


15-D is much better than Halton.
```{r}
npoints = 50
xy = sobol(npoints,dim=15)
qmc <- data.frame(j = 1:npoints)
qmc[,2:16] = xy
ggpairs(qmc,columns=10:15)
```

#Torus

In 1-D the torus is the fractional part of $\sqrt{2}$:
```{r}
npoints = 10
qmc <- data.frame(j = 1:npoints, x = torus(npoints), y = (1:npoints)*sqrt(2))
ggplot(qmc,aes(x=x,y=y,color=j)) + geom_point(size=4) + scale_y_continuous('j times root 2') + scale_colour_gradient(low="blue",high="red")
```

In 2-D the torus doesn't appear too random:
```{r}
npoints = 50
xy = torus(npoints,dim=2)
qmc <- data.frame(j = 1:npoints)
qmc[,c("x","y")] = xy
ggplot(qmc,aes(x=x,y=y)) + geom_point(size=4)
```

In 6-D torus is probably the worst
```{r}
npoints = 50
xy = torus(npoints,dim=6)
qmc <- data.frame(j = 1:npoints)
qmc[,2:7] = xy
ggpairs(qmc,columns=2:7)
```


15-D
```{r}
npoints = 50
xy = torus(npoints,dim=15)
qmc <- data.frame(j = 1:npoints)
qmc[,2:16] = xy
ggpairs(qmc,columns=10:15)
```
